---
title: "CAWR Gut Contents"
author: "cnell"
date: "June 14, 2016"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r,loadpackage,echo=F,message=F,warning=F}
library(dplyr)
library(ggplot2)
library(reshape2)
library(pander)
library(shiny)
library(DT)
library(plotly)
```

```{r,loaddata,echo=F,message=F}
ids<-read.csv("/Users/colleennell/Documents/R/CAWR/CAWR_OTU_finalids.csv")###this is including bad matches 'unknown' ids
samples<-read.csv("/Users/colleennell/Documents/R/CAWR/data/CAWR_samples.csv")##sample data frame
guts<-read.csv("/Users/colleennell/Documents/R/CAWR_sample_ids.csv")##this is the above 2 combined with a few more cols

```

```{r,predataprep,echo=F,message=F}
arth.matrix<-select(guts,starts_with("S"))
taxa.ids<-select(guts,starts_with("Taxa."))


```

##**ID Results**  

###Process  
  1. Search sequence in BOLD (http://www.boldsystems.org/index.php/IDS_OpenIdEngine)  
  2. Record top similarity match  
  3. If matches with multiple IDs from different orders/families, search NCBI   
  4. Look for consensus between databases  
  5. For mixed results, choose conservative grouping for ID  
  6. Unclear still, compare results to phylogenetic tree  
  7. Delete OTU's with poor matches (below 85%, unable to determine order, non arthropod)   8. Rebuild tree and check IDs against  
  
  Taxonomic ID Resolution (Zeale et al, 2011):  
  Species = 99.3%; Genus = 94.9%; Family = 91%; Order = 85.9%  
    
**Mean % similarity in BOLD database: `r round(mean(ids$Similarity.BOLD ,na.rm=T),digits=2)`  **

  _However it should be considered whether these are conservative enough, some literature suggests >97.8% match to BOLD database_

  
###Taxonomy Table  
'unknown' IDs were < 85.9% similarity to the database or conflicitng IDs at the order-level    
```{r,filterids,echo=F,message=F}
realid<-ids%>%
  filter(Taxa.ID!='unknown')

```
`r length(realid$Taxa.ID)`  OTUs    
`r length(unique(realid$Taxa.ID))` Unique taxa IDs   
`r length(unique(realid$Taxa.Order))` Arthropod Orders: `r unique(realid$Taxa.Order)`   

```{r, idtable,echo=FALSE,results='asis'}
ids$Taxa.Genus<-as.factor(ifelse(ids$reso=='genus',paste(ids$Taxa.ID),
                       ifelse(ids$reso=='species',paste(ids$GENUS),NA )))
ids$Taxa.Sp<-as.factor(ifelse(ids$reso=='species',paste(ids$Taxa.ID),NA ))
ids$Taxa.Family<-ifelse(ids$Taxa.Family=='unknown'| ids$Taxa.Family=='None',NA,paste(ids$Taxa.Family))

id.table<-ids%>%
  group_by(Taxa.ID,Taxa.Order,Taxa.Family,Taxa.Genus,Taxa.Sp)%>%
  summarize(n_otu=length(OTU))

taxon.table<-DT::datatable(id.table,extensions='Buttons',rownames=F,style='bootstrap',
                           colnames=c('ID','Order','Family','Genus','Species','#OTUs'),
                           options=list(order=list(5,'dsc'),scrollY="350px",scrollX=F,paging=F,dom='Bft',buttons=list('csv','copy')))
taxon.table


```
  
###Sample Composition  

```{r,idsbysample,echo=F,message=F,warning=F}
sample.matrix<-select(guts,starts_with("S"))
order<-guts$Taxa.Order
sample.matrix<-cbind(order,sample.matrix)
melted.mat<-melt(sample.matrix)
melted.mat$value<-ifelse(melted.mat$value>0,1,0)
melted.mat<-melted.mat%>%
  group_by(order,variable)%>%
  summarize(n_samps = sum(value))
sample.matris.t<-dcast(melted.mat,variable~order)
samps<-sample.matris.t[,1]
rowsum<-rowSums(sample.matris.t[,-1])
prop.mat<-mutate_each(sample.matris.t[,-1],funs(./rowsum*100))
prop.mat<-cbind(samps,prop.mat)
neword<-melt(prop.mat)

x<-list(title = "Sample")
y<-list(title="Arthropod Composition \n(% of OTUs)")
m = list(
  l = 70,
  r = 50,
  b = 70,
  t = 50,
  pad = 14
)

samp.comp<-ggplot(neword,aes(x=samps,y=value,fill=variable))+
  geom_bar(color="black",stat="identity")+coord_flip()+theme_minimal()+labs(x=" ",y=" ")+theme(panel.background = element_rect(fill = "transparent",colour = NA))+scale_fill_brewer(name="Order",palette = 'Spectral')
                       
s<-ggplotly(samp.comp)
s2<-layout(s,xaxis=y,yaxis=x,autosize=F,width = 700, height = 700, margin = m)
s2
```
  
      
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
 <br>
<br>
<br>
<br>           
              
               
               
        
        
###Sample Summary

```{r,avgsamp,warning=F,message=F,echo=F}
std <- function(x) sd(x)/sqrt(length(x))

newsamp<-neword%>%
  group_by(variable)%>%
  summarize(avg.per = mean(value), se = std(value))

x<-list(title = "Order")
y<-list(title="Average Proportion of Diet")
m = list(
  l = 70,
  r = 50,
  b = 120,
  t = 50,
  pad = 14
)
avgplot<-ggplot(newsamp,aes(x=variable,y=avg.per,fill=variable))+
  geom_bar(color="black",stat="identity")+
  geom_errorbar(aes(ymin=avg.per-se,ymax=avg.per+se),width=.2)+
  theme_minimal()+
  theme(panel.background = element_rect(fill = "transparent",colour = NA),axis.text.x=element_text(angle=45))+
  scale_fill_brewer(name="Order",palette = 'Spectral')+
  labs(x=" ",y=" ")
                                                                                           
avgplots<-ggplotly(avgplot)
t3<-layout(avgplots,xaxis=x,yaxis=y,autosize=F,width = 700, height = 500, margin = m)
t3

```



